### 2015-07-10 narrative/data-description.tex
figure with R instead of F, and I need Rutgers as a separate institution
trying to distinguish now between value and priority.

The data description may be stored with the individual chunks or within a
separate metadata service or both to aid QoS requirements. It would contain
conventional attributes such as data type, size, dimensionality, and the
relationship to other data. In particular, the relationship to other chunks can
be captured by including \textit{typed links}, each representing a different
relationship type.  In this project, a key metric stored as a metadata
attribute is \textit{data utility}.  It enables QoS scheduling, data placement
decisions, data lifetime decisions, and captures the chunk priority. It is
our belief that exascale science will require users to prioritize a small set
of chunks to be saved on higher-level capacity-limited storage to avoid the
slow access to large-capacity lower-level storage, such as tape.  We propose a
new technique, described by a utility function provided by a user, where chunks
can be cast into multiple data buckets and each bucket can be prioritized
differently according to its utility value.  Once this is done, the middleware
system will construct the data description and re-organize and place data to
achieve the desired QoS goals and policies.

@@

From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1~\cite{chang2006integrated}, GTC~\cite{klasky2003grid} , and SPECFM3D~\cite{SPECFEM3D}. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
\S~\ref{sec:data-refactor}) which will classify the particle data into four different
prioritization levels. The user can then pick a data utility, such as three months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
% Description of the utility function research

@@

\begin{wrapfigure}{r}{0.6\textwidth}
        \begin{centering} 
	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
        %\vspace{1ex}
        \caption{The system manages the data across the various stages of the data lifecycle}
        \label{fig:ssio-bucket}
        \end{centering}
      % \vspace{-1ex}
\end{wrapfigure}

@@

 We will formulate the concept of utility (from an
application perspective) that is assigned data objects to quantify the
benefits of placement and/or movement actions to application. These
utilities may be based on the importance of the data objects, the frequency
with which they are accessed, etc., and can be used during runtime data
management to evaluate and compare different placement/movement options. 
For example, data objects with higher utility value may be
placed closer (e.g., in faster memory) to the accessing application.
Similarly, when it is time to evict data objects from a higher level storage
layer, the decision about which object to evict may take the utility into
account. Our initial work in defining such a utility is presented in~\cite{tongipdps15}.
 We will also explore autonomic approaches for balancing of the placement and 
 movement of data across the memory hierarchy autonomically, which will be driven 
 by the data utility and will leverage both user hints and information gathered at runtime (e.g., runtime data access
history, network topology, etc.).
Finally, we will explore a complementary notion of utility from a system-wide perspective 
that will capture and quantify, for example, the impact of placement decisions on 
the overall utilization of storage across multiple applications, and will enable system level runtime optimization.

@@

We will design new techniques, available to both users and the storage system,
to raise or lower these sub-chunks up and
down the storage stack. The SSIO layer will then manage the system metadata so
that the user can transparently access any of these sub-chunks without
the need to know where the data currently resides.
At read time, the middleware will be required to 
provide a time estimate, and then allow the users to decide if
they will read less (e.g., only the highest priority sub-chunks)
if the time estimate
for reading does not meet requirements. Existing Sirocco functionality may be
sufficient for these purposes, but it must still be proven in practice.
%

>>>>

The data description may be stored with the individual chunks or within a
separate metadata service or both to aid QoS requirements. It would contain
conventional attributes such as data type, size, dimensionality, and the
relationship to other data. In particular, the relationship to other chunks can
be captured by including \textit{typed links}, each representing a different
relationship type.  In this project, a key metric stored as a metadata
attribute is \textit{data utility}.  It enables QoS scheduling, data placement
decisions, data lifetime decisions, and captures the value of the chunk. It is
our belief that exascale science will require users to prioritize a small set
of chunks to be saved on higher-level capacity-limited storage to avoid the
slow access to large-capacity lower-level storage, such as tape.  We propose a
new technique, described by a utility function provided by a user, where chunks
can be cast into multiple data buckets and each bucket can be prioritized
differently according to its utility value.  Once this is done, the middleware
system will construct the data description and re-organize and place data to
achieve the desired QoS goals and policies.

@@

From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1~\cite{chang2006integrated}, GTC~\cite{klasky2003grid} , and SPECFM3D~\cite{SPECFEM3D}. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
\S~\ref{sec:data-refactor}) which will classify the particle data into four different
value levels. The user can then pick a data utility, such as three months
for the highest value items and 12 months for the lowest value data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest value element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest value buckets
will be stored in the Parallel File System, but the next highest value items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
% Description of the utility function research

@@

@@

 We will formulate the concept of utility (from an
application perspective) that is assigned data objects to quantify the
benefits of placement and/or movement actions to application. These
utilities may be based on the importance of the data objects, the frequency
with which they are accessed, etc., and can be used during runtime data
management to evaluate and compare different placement/movement options. 
For example, data objects with higher utility value may be
placed closer (e.g., in faster memory) to the accessing application.
Similarly, when it is time to evict data objects from a higher level storage
layer, the decision about which object to evict may take the utility into
account. Our initial work in defining such a utility is presented in~\cite{tongipdps15}.
 We will also explore autonomic approaches for balancing of the placement and 
 movement of data across the memory hierarchy autonomically, which will be driven 
 by the data utility and will leverage both user hints and information gathered at runtime (e.g., runtime data access
history, network topology, etc.).
Finally, we will explore a complementary notion of utility from a system-wide perspective 
that will capture and quantify, for example, the impact of placement decisions on 
the overall utilization of storage across multiple applications, and will enable system level runtime optimization. 
\begin{wrapfigure}{R}{0.6\textwidth}
        \begin{centering} 
	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
        %\vspace{1ex}
        \caption{The system manages the data across the various stages of the data lifecycle}
        \label{fig:ssio-bucket}
        \end{centering}
      % \vspace{-1ex}
\end{wrapfigure}

@@

We will design new techniques, available to both users and the storage system,
to raise or lower these sub-chunks up and
down the storage stack. The SSIO layer will then manage the system metadata so
that the user can transparently access any of these sub-chunks without
the need to know where the data currently resides.
At read time, the middleware will be required to 
provide a time estimate, and then allow the users to decide if
they will read less (e.g., only the highest value sub-chunks)
if the time estimate
for reading does not meet requirements. Existing Sirocco functionality may be
sufficient for these purposes, but it must still be proven in practice.
%

### 2015-07-10 narrative/evaluation.tex
reference fixes
changed the acronym.
put table title on top.
trying to distinguish now between value and priority.

\centering
\begin{tabular}{ | p{1.4in} | p{2in} | p{2.5in} | }
  \hline
  {\bf Principle/Metric}                   & {\bf User perspective metric}   & {\bf System perspective metric }\\ \hline
  {\bf Knowledge-centric system}           & More knowledge = faster access to higher quality data &  More knowledge = better resource utilization across the entire hierarchy \\ \hline
  {\bf Predictable Performance Accuracy}   & More predictability = better quality of data &  Accuracy of the time bounds across multiple layers of the storage hierarchy wrote/read in time constrained situations\\ \hline
\end{tabular}

\caption{Guiding Principles}
\label{table:eval}
\end{table}

In addition to benchmarking the various S2E2 methods and tools in isolation,
we will work with our science partners at Princeton University (J. Tromp),
PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
to evaluate the leading edge applications working with the storage system to
determine performance predictability and data quality.  Our evaluation 
will use the multiple storage layers available on systems at both SNL and the
OLCF.  For example, the sith cluster at the OLCF
(https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
storage system, SSDs on each node, and is connected to the center-wide Lustre
file system. We will perform tests in both the system level (on sith) and in
the user level (on titan), and evaluate our results using our application
knowledge.

@@

The fifth part of the evalation will evaluate how well additional data
knowledge in the system will lead to more performance by placing the most
important pieces at the higher layers of the hierarchy. We will test different
knowledge levels and evaluate the system impact.  For example, if the user
doesn't want the data broken into priority chunks, then the data will be at
the mercy of what the system does when there is no knowledge. If the user then 
provides sufficient information to use ``generic'' rouitnes based on frequency
or accuracy, then that knowledge will be used to decompose the objects into
specially placed chunks. We will evaluate these no and minimal knowledge cases
against those that provide additional knowledge allowing application level data
binning (for C.S. Chang) or for level of features (for J. Tromp).

>>>>

\centering
\caption{Guiding Principles}
\label{table:eval}
\begin{tabular}{ | p{1.4in} | p{2in} | p{2.5in} | }
  \hline
  {\bf Principle/Metric}                   & {\bf User perspective metric}   & {\bf System perspective metric }\\ \hline
  {\bf Knowledge-centric system}           & More knowledge = faster access to higher quality data &  More knowledge = better resource utilization across the entire hierarchy \\ \hline
  {\bf Predictable Performance Accuracy}   & More predictability = better quality of data &  Accuracy of the time bounds across multiple layers of the storage hierarchy wrote/read in time constrained situations\\ \hline
\end{tabular}
\end{table}

In addition to benchmarking the various SIRIUS methods and tools in isolation,
we will work with our science partners at Princeton University (J. Tromp),
PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
to evaluate the leading edge applications working with the storage system to
determine performance predictability and data quality.  Our evaluation 
will use the multiple storage layers available on systems at both SNL and the
OLCF.  For example, the sith cluster at the OLCF
(https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
storage system, SSDs on each node, and is connected to the center-wide Lustre
file system. We will perform tests in both the system level (on sith) and in
the user level (on titan), and evaluate our results using our application
knowledge.

@@

The fifth part of the evaluation will evaluate how well additional data
knowledge in the system will lead to more performance by placing the most
important pieces at the higher layers of the hierarchy. We will test different
knowledge levels and evaluate the system impact.  For example, if the user
doesn't want the data broken into {\it value} chunks, then the data will be at
the mercy of what the system does when there is no knowledge. If the user then 
provides sufficient information to use ``generic'' routines based on frequency
or accuracy, then that knowledge will be used to decompose the objects into
specially placed chunks. We will evaluate these no and minimal knowledge cases
against those that provide additional knowledge allowing application level data
binning (for C.S. Chang) or for level of features (for J. Tromp).

### 2015-07-10 narrative/placement.tex
trying to distinguish now between value and priority.

\subsubsection*{Making data available on multiple tiers}
\label{sec:manag-data-mult}
 Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
fully accurate data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \S\ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
high priority  data sets on the higher tiers for faster access.

>>>>

\subsubsection*{Making data available on multiple tiers}
\label{sec:manag-data-mult}
 Providing data to users for knowledge discovery is
one of the primary purposes of our storage system design. There is a
spectrum of approaches on how
this can be accomplished: On one end the storage system can aim to maintain
fully accurate data in the highest tiers of the storage hierarchy. This
approach is, obviously, limited due to the disparity between size of data
and available space in the higher tiers. Another approach is to utilize the
techniques described in \S\ref{sec:data-refactor} to create auxiliary views of
the data that, with user defined error, represent the data, and place these
{\it high value}  data chunks on the higher tiers for faster access.

### 2015-07-10 narrative/proposed-work.tex
trying to distinguish now between value and priority.

One of the underlying principles of our approach is that we recognize that there are a small set
of important I/O patterns to optimize for~\cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
and our representative applications have worked with us on all of these I/O patterns. Throughout our project
we will continuously focus on creating research artifacts which can optimize these patterns on current generation through
exascale systems.  One important pattern is Checkpoint/Restarts (CR) and we will focus on optimizing these through our
research, and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing 
the CR data. The migration and purging of this will be optimized by our SSIO layer, and we will have the user data intentions
signify the utility of CR data so that this data can be high-priority but stored for short periods of time.

>>>>

One of the underlying principles of our approach is that we recognize that there are a small set
of important I/O patterns to optimize for~\cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
and our representative applications have worked with us on all of these I/O patterns. Throughout our project
we will continuously focus on creating research artifacts which can optimize these patterns on current generation through
exascale systems.  One important pattern is Checkpoint/Restarts (CR) and we will focus on optimizing these through our
research, and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing 
the CR data. The migration and purging of this will be optimized by our SSIO layer, and we will have the user data intentions
signify the utility of CR data so that this data can be high-value but stored for short periods of time. The value of a chunk of data
from an object signifies it's importance to the users relative to the other chunks of their data.

### 2015-07-10 narrative/qos.tex
trying to distinguish now between value and priority.

As stated in the introduction, we seek to make a storage stack that works with
the user. Towards that end, we must incorporate data attributes into this
process. For example, particularly for a capability run, the storage system
should give priority to data subset tagged with high importance to stay in fast
storage tiers. Enabling this priority will be driven by other job
characteristics such as job size or a manually set priority inserted by system
administrators. It should also store the highest quality data possible given
the performance requirement. Ideally, the storage system will store the raw
data at full fidelity. When this tier approaches capacity, data should migrate
not just based on age, but also based on this priority. This will help ensure
the shortest time to insight by making sure the most important machine runs
have priority to use system resources.

\paragraph{Related Work:}
On current production HPC systems, absolute I/O performance guarantees are not
possible. Previous work~\cite{lofstead:2010:io-variability,liu_hotstorage} from
this project team uses a client-assisted approach to mitigate I/O variablity at
scale (100,000-core).  Recently, server side QoS scheduling has also been
studied for HPC applications~\cite{Dai:2014}.  However, the key hurdle of
scalability has not been solved.  Existing efforts on scheduing
resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not require
modification to use new storage access APIs. By using a priority tagging
system, we will offer defaults suitable for all applications that can be
informed by simple extensions to the job scheduler or more advanced management
through new APIs. On the other hand, storage QoS has been studied for
enterprise applications and systems~\cite{Gulati:2007,Gulati:2010,Gulati:2012}
in virtualized environments. It is not straightforward to adapt these solutions
to HPC storage systems because of the scalability and efficiency concerns,
despite they achieved very good performance isolations.

>>>>

As stated in the introduction, we seek to make a storage stack that works with
the user. Towards that end, we must incorporate data attributes into this
process. For example, particularly for a capability run, the storage system
should give (on a user by user basis), higher priority to data subset tagged with high importance to stay in fast
storage tiers.
% Enabling this priority will be driven by other job
%characteristics such as job size or a manually set priority inserted by system
administrators.
 It should also store the highest value data possible given
the performance requirement. Ideally, the storage system will store the raw
data at full fidelity. When this tier approaches capacity, data should migrate
not just based on age, but also based on this value. One area of research will be in
trying to see if we can retain some guarantees of QoS while giving some high-profile users
greater SSIO a higher {\bf priority} than other users on the system. We realize that this makes
it very difficult to main overall system QoS, but we will investigate techniques when one
user is running on the majority of the exascale resource and we can allow them to lock out other SSIO
request during small periods of time during the lifetime of their exascale simulation.

\paragraph{Related Work:}
On current production HPC systems, absolute I/O performance guarantees are not
possible. Previous work~\cite{lofstead:2010:io-variability,liu_hotstorage} from
this project team uses a client-assisted approach to mitigate I/O variablity at
scale (100,000-core).  Recently, server side QoS scheduling has also been
studied for HPC applications~\cite{Dai:2014}.  However, the key hurdle of
scalability has not been solved.  Existing efforts on scheduing
resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not require
modification to use new storage access APIs. By using a {\it value} tagging
system, we will offer defaults suitable for all applications that can be
informed by simple extensions to the job scheduler or more advanced management
through new APIs. On the other hand, storage QoS has been studied for
enterprise applications and systems~\cite{Gulati:2007,Gulati:2010,Gulati:2012}
in virtualized environments. It is not straightforward to adapt these solutions
to HPC storage systems because of the scalability and efficiency concerns,
despite they achieved very good performance isolations.
